\thetitle

### 2021-06-27 writing/challenges.tex
Small grammar + typo fixes, and minor wording changes

One way to reconcile these differences is to {\em equip the rational
programmer with modes \/} that represent the different types of
information the rational programmer takes into account when debugging a
scenario. Intuitively, different
blame strategies correspond to different modes of operation.
For instance, one Transient mode may assign types to the oldest element of a
blame sequence because it corresponds to the
earliest point in the execution that can discover an impedance
mismatch.  Another mode may opt to treat the sequence as a stack and add
types to its newest element.  If both modes are equally successful in
locating an impedance mismatch, measuring the rational programmer's effort with each mode
may answer which is the most effective.

>>>>

One way to reconcile these differences is to {\em equip the rational
programmer with modes \/} that represent the different types of
information the rational programmer takes into account when debugging a
scenario. Intuitively, different
blame strategies correspond to different modes of operation.
For instance, one Transient mode may assign types to the oldest element of a
blame sequence because it corresponds to the
earliest point in the execution that can discover an impedance
mismatch.  Another mode may opt to treat the sequence as a stack and add
types to its newest element.  If both modes are equally successful in
locating an impedance mismatch, measuring the rational programmer's debugging effort with each mode
may answer which is the most effective.

### 2021-06-27 writing/landscape.tex
Small grammar + typo fixes, and minor wording changes

The mistaken comment in {\tt pack-lib} causes an impedance mismatch, with which each
of the three semantics deals differently.  Under {\it the Natural semantics\/},
functions imported into and exported from typed modules are wrapped in proxies that
enforce the static type discipline with run-time checks and track
responsibilities~\citep{tf-popl-2008, tfffgksst-snapl-2017}. Thus, when {\tt pack}
is imported into a typed module, the run-time system checks that it is a function
and wraps it in a protective proxy, which in turn, enforces the type of the function
result with run-time checks.  Analogously the run-time system wraps each exported
function of a typed module such as {\tt crypto-pack} in a proxy that checks its
arguments.  These checks protect functions exported from typed modules against
applications to wrong arguments in untyped code.

As this analysis implies, if a return-type check fails, the problem is
that the untyped module, here {\tt pack-lib}, supplied a function that is not a
match for the type ascribed by the typed module.
Hence either the type at the boundary between the two modules is wrong or, 
if the programmer trusts the type, the untyped module is at fault.
If an argument-type check fails, responsibility  lies with
{\tt client}. After all, either the type it ascribes to the argument is
wrong or the argument it produced clashes with the type. Due to 
proxies, {\it Natural\/} can easily track the 
boundary, type, and responsible parties that correspond to each check. 
Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary between \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypto-pack-lib}, respectively, for the two {\tt define}s in {\tt client}.

Under {\it Transient\/}, typed code is compiled so that all entry points to
functions check their arguments at run time and all function calls check their
return values against the expected type~\citep{vss-popl-2017}.  Furthermore,
Transient uses \emph{shallow} checks, meaning they inspect only the top
constructor of a value. Since retrieving a value from within a structure (or
list, array, hash table etc.) is performed via a function call, the contents of
a complex value is checked on a piecemeal basis.

>>>>

The mistaken comment in {\tt pack-lib} causes an impedance mismatch, with which each
of the three semantics deals differently.  Under {\it the Natural semantics\/},
functions imported into and exported from typed modules are wrapped in proxies that
enforce the static type discipline with run-time checks and track
responsibilities~\citep{tf-popl-2008, tfffgksst-snapl-2017}. Thus, when {\tt pack}
is imported into a typed module, the run-time system checks that it is a function
and wraps it in a protective proxy, which in turn enforces the type of the function
result with run-time checks.  Analogously, the run-time system wraps each exported
function of a typed module such as {\tt crypto-pack} in a proxy that checks its
arguments.  These checks protect functions exported from typed modules against
applications to wrong arguments in untyped code.

As this analysis implies, if a return-type check fails, the problem is
that the untyped module, here {\tt pack-lib}, supplied a function that is not a
match for the type ascribed by the typed module.
Hence either the type at the boundary between the two modules is wrong or, 
if the programmer trusts the type, the untyped module is at fault.
If an argument-type check fails, responsibility  lies with
{\tt client}. After all, either the type it ascribes to the argument is
wrong or the argument it produced clashes with the type. Due to 
proxies, {\it Natural\/} can easily track the 
boundary, type, and responsible parties that correspond to each check. 
Thus, in the example of figure~\ref{fig:example}, as
\texttt{pack} returns, the return-type check fails and Natural blames
the boundary of \texttt{pack-lib} with {\tt typed-pack-lib} and
{\tt crypto-pack-lib}, respectively, for the two {\tt define}s in {\tt client}.

Under {\it Transient\/}, typed code is compiled so that all entry points to
functions check their arguments at run time and all function calls check their
return values against the expected type~\citep{vss-popl-2017}.  Furthermore,
Transient uses \emph{shallow} checks, meaning they inspect only the top
constructor of a value. Since retrieving a value from within a structure (or
list, array, hash table etc.) is performed via a function call, the contents of
a complex value are checked on a piecemeal basis.

### 2021-06-27 writing/rational-transient.tex
Small grammar + typo fixes, and minor wording changes

The definition of \emph{Transient exception trails} is analogous to that for
Natural. It is used to as a baseline for Transient-first and Transient-last
blame.\footnote{A reader may wonder whether the rational programmer
should not just equip {\em all\/} modules in the Transient blame set
with types. Doing so may accelerate the search for the impedance
mismatch but it would also impose a large migration cost.}

>>>>

The definition of \emph{Transient exception trails} is analogous to that for
Natural. It is used as a baseline for Transient-first and Transient-last
blame.\footnote{A reader may wonder whether the rational programmer
should just equip {\em all\/} modules in the Transient blame set
with types. That might accelerate the search for the impedance
mismatch, but if so it would also impose a large migration cost.}

### 2021-06-27 writing/why-rational.tex
Small grammar + typo fixes, and minor wording changes

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer-typed)}. As
predicted by the theory, running the modified program (in the same way as before) yields a different error message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
Lastly, the rational programmer assigns types to {\tt server} and to re-direct
the import of {\tt layer-typed} to {\tt (submod ".." server-typed)}. Now the
type checker objects to the conjectured type of {\tt neg-abs}, i.e.\ the source
of the impedance mismatch is found. How to fix it is a separate question.

>>>>

Assuming the rational programmer trusts the type of {\tt na-client}, the next
step is to inspect the {\tt layer} module and to equip it with type
annotations. The result is the \typecolor\ module in the middle, and {\tt
main}'s import is now re-directed there by {\tt (submod ".." layer-typed)}. As
predicted by the theory, running the modified program (in the same way as before) yields a different error message:
\begin{verbatim}
    neg-abs: broke its own contract
      promised: (<=/c 0)
      produced: 10
      in: (-> any/c (<=/c 0))
      contract from: (interface for neg-abs)
      blaming: (interface for neg-abs)
       (assuming the contract is correct)
\end{verbatim}
Lastly, the rational programmer assigns types to {\tt server} and re-directs
the import of {\tt layer-typed} to {\tt (submod ".." server-typed)}. Now the
type checker objects to the conjectured type of {\tt neg-abs}, i.e.\ the source
of the impedance mismatch is found. How to fix it is a separate question.
